### 2012-06-14 inst/examples/value_of_information.Rmd
value of information now handles all cases

`ro cache=FALSE, tidy=FALSE, warning=FALSE, comment=NA, message=FALSE, fig.width=10, fig.height=10, cache.path="value_of_information/", verbose=TRUE or`

@@

# Calculating the value of information
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

 Implements a numerical version of the SDP described in `r knitcitations::citep("10.1016/j.jeem.2004.11.005")`.
 Compute the optimal solution under different forms of uncertainty.  The true uncertainty is much smaller than that assumed, just a small uncertainty in growth.  We compare the case where the uncertainty is equal to this tiny growth uncertainty to policies derived under much greater uncertainty assumptions, and see how the resulting profit varies.

``` {r  setup, echo=FALSE, cache=FALSE}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

Define noise parameters

Chose the state equation / population dynamics function

@@

Note that the `pdg_control` package already has a definition for the `BevHolt` function, (typing the function name prints the function)

``` {r }
BevHolt
````

That is, \\( f(x,h) = \frac{A x}{1 + B x} \\)

Of course we could pass in any custom function of stocksize `x`, harvest `h` and parameter vector `p` in place of `BevHolt`.  Note that we would need to write this function explicitly so that it can take vector values of `x` (i.e. uses `sapply`), an annoying feature of `R` for users coming from Matlab.

We must now define parameters for the function.  Note that the positive stationary root of the model is given by \\( \frac{A-1}{B} \\), which we'll store for future reference as `K`.

@@

and we use a harvest-based profit function with default parameters

``` {r  profit}
profit <- profit_harvest(price=1, c0 = 0.01) 
````

The `profit_harvest` function has the form \\( \Pi = h - \left( c_0  + c_1 \frac{h}{x} \right) \frac{h}{x} \\), conditioned on \\( h > x \\) and \\(x > 0 \\).  Note that the R code defines a function from another function using a trick known as a _closure_.  Again we could write a custom profit function as long as it can take a vector stock size `x` and a scalar harvest level `h`.  Details for provided functions can be found in the manual, i.e. `?profit_harvest`.

Now we must set up the discrete grids for stock size and havest levels (which will use same resolution as for stock), in order to calculate the SDP solution.   Here we set the gridsize to 100.

``` {r  grid}
x_grid <- seq(0, 2 * K, length = 100)  
h_grid <- x_grid  
````

@@

We calculate the stochastic transition matrix for the probability of going from any state \\(x_t \\) to any other state \\(x_{t+1}\\) the following year, for each possible choice of harvest \\( h_t \\).  This provides a look-up table for the dynamic programming calculations.

In the Sethi case, computing the distribution over multiple sources of noise is actually quite difficult.  Simulation turns out to be more efficient than numerically integrating over each distribution.

## No Uncertainty

``` {r noisefns}
sigma_g <- 0.01    # Noise in population growth
sigma_m <- 0.0     # noise in stock assessment measurement
sigma_i <- 0.0     # noise in implementation of the quota
z_g <- function() 1 
z_m <- function() 1
z_i <- function() 1 
````
Find the transition matrix

``` {r  SDP_Mat}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
````

Find the optimum solution

``` {r  find_dp_opt }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate

``` {r  simulate }
sims_known <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

## Growth uncertainty

``` {r }
sigma_g <- 0.15    # Noise in population growth
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
````
Find the transition matrix

``` {r }
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
````

Find the optimum solution

``` {r }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate.  The real world doesn't have as much uncertainty as we do in our knowledge of the state.

``` {r }
sigma_g <- 0.05    # Noise in population growth
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
sims_g <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, profit=profit)
})
````

## Growth & stock measurement uncertainty

``` {r }
sigma_g <- 0.15    # Noise in population growth
sigma_m <- 0.15     # noise in stock assessment measurement
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() 1
````

Find the transition matrix.  Use the simulation method to account for the extra uncertainties

``` {r }
require(snowfall) 
sfInit(parallel=TRUE, cpu=16)
SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
````

Find the optimum solution

``` {r }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate

``` {r }
sigma_g <- 0.05    # Noise in population growth
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() 1
z_i <- function() 1
sims_gm <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

## Growth, stock measurement & implementation uncertainty

``` {r }
sigma_g <- 0.15    # Noise in population growth
sigma_m <- 0.15     # noise in stock assessment measurement
sigma_i <- 0.15     # noise in implementation of the quota
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() rlnorm(1,  0, sigma_m) # mean 1
z_i <- function() rlnorm(1,  0, sigma_i) # mean 1
````

Find the transition matrix.  Use the simulation method to account for the extra uncertainties

``` {r }
SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
````

Find the optimum solution

``` {r }
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

Simulate

``` {r }
sigma_g <- 0.05    # Noise in population growth
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() 1
z_i <- function() 1
sims_gmi <- lapply(1:100, function(i){
  ForwardSimulate(f, pars, x_grid, h_grid, x0=K, opt$D, z_g, z_m, z_i, profit)
})
````

## Summarize and plot the results

R makes it easy to work with this big replicate data set.  We make data tidy (melt), fast (data.tables), and nicely labeled.

``` {r  tidy}

sims <- list(known = sims_known, growth = sims_g, growth_stock = sims_gm, growth_stock_harvest = sims_gmi)

dat <- melt(sims, id=names(sims_known[[1]]))  
dt <- data.table(dat)
setnames(dt, c("L2", "L1"), c("reps", "uncertainty")) # names are nice
````

### Plots

Let's begin by looking at the dynamics of a single replicate. The line shows Reed's S, the level above which the stock should be harvested (where catch should be the difference between stock and S).  To confirm that this policy is being followed, note that harvesting only occurs when the stock is above this line, and harvest is proportional to the amount by which it is above.  Change the replicate `reps==` to see the results from a different replicate.

``` {r  onerep }
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, fishstock)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_line(aes(time, harvest), col="darkgreen") + 
  facet_wrap(~uncertainty) 
````

This plot summarizes the stock dynamics by visualizing the replicates. Reed's S shown again.

``` {r  all}
p1 <- ggplot(dt) + geom_abline(intercept=opt$S, slope = 0) 
p1 + geom_line(aes(time, fishstock, group = reps), alpha = 0.2) + facet_wrap(~uncertainty)
````

We can also look at the harvest dynamics:

``` {r  harvestplot}
p1 + geom_line(aes(time, harvest, group = reps), alpha = 0.1, col="darkgreen") + facet_wrap(~uncertainty)
````

This strategy is supposed to be a constant-escapement strategy. We can visualize the escapement:

``` {r  escapement}
p1 + geom_line(aes(time, escapement, group = reps), alpha = 0.1, col="darkgrey") + facet_wrap(~uncertainty)
````

``` {r }
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, profit))  + facet_wrap(~uncertainty)

````

``` {r }
profits <-dt[ , sum(profit), by=c("reps", "uncertainty")] 
ggplot(profits) + geom_histogram(aes(V1)) + facet_wrap(~uncertainty)
````

Summary stats

``` {r }
profits[, mean(V1), by=uncertainty]
profits[, sd(V1), by=uncertainty]
````

>>>>

`ro cache=TRUE, warning=FALSE, comment=NA, message=FALSE, cache.path="value_of_information/", verbose=TRUE or`

@@

# Calculating the value of information

 Implements a numerical version of the SDP described in `r knitcitations::citep("10.1016/j.jeem.2004.11.005")`.
 Compute the optimal solution under different forms of uncertainty.

``` {r  setup, echo=FALSE, cache=FALSE}
# load required libraries
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

We consider a Beverton Holt state equation governing population dynamics, \\( f(x,h) = \frac{A x}{1 + B x} \\)

@@

With parameters `A` = `r A <- 1.5` and `B` = `r B <- 0.05`.

@@

Note that the positive stationary root of the model is given by \\( \frac{A-1}{B} \\), or carring capacity `K` = `r K`.  
We consider a profits from fishing to be a function of harvest `h` and stock size `x`,  \\( \Pi(x,h) = h - \left( c_0  + c_1 \frac{h}{x} \right) \frac{h}{x} \\), conditioned on \\( h > x \\) and \\(x > 0 \\),

``` {r  profit}
price <- 1
c0 <- 0.01
c1 <- 0
profit <- profit_harvest(price=price, c0 = c0, c1=c1) 
````

with price `r price`, `c0` `r c0` and `c1` `r c1`.

``` {r }
xmin <- 0
xmax <- 1.5 * K
grid_n <- 100
````

We seek a harvest policy which maximizes the discounted profit from the fishery using a stochastic dynamic programming approach over a discrete grid of stock sizes from `r xmin` to `r xmax` on a grid of `r grid_n` points, and over an identical discrete grid of possible harvest values.

``` {r  grid}
x_grid <- seq(xmin, xmax, length = grid_n)  
h_grid <- x_grid  
````

``` {r pars}
delta <- 0.05
xT <- 0
OptTime <- 25
````

We will determine the optimal solution over a `r OptTime` time step window with boundary condition for stock at `r xT` and discounting rate of `r delta`.

@@

We use Monte Carlo integration over the noise processes to determine the transition matrix.

```{r beparallel, cache=FALSE}
sfInit(parallel=TRUE, cpu=16)
sfExportAll()
````

``` {r policyscenarios}
scenario <- function(policy_g, policy_m, policy_i){

  z_g <- function() rlnorm(1,  0, policy_g) 
  z_m <- function() rlnorm(1,  0, policy_m) 
  z_i <- function() rlnorm(1,  0, policy_i)

  SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=2e4)
  opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                       profit, delta, reward=0)
}
````

``` {r runall}
det <- scenario(0, 0.0, 0.0)
g <- scenario(0.2, 0.0, 0.0)
m <- scenario(0.0, 0.2, 0.0)
i <- scenario(0.0, 0.0, 0.2)
gm <- scenario(0.2, 0.2, 0.0)
gi <- scenario(0.2, 0.0, 0.2)
gmi <- scenario(0.2, 0.2, 0.2)
````

### plots

``` {r }
require(reshape2)
policy <- melt(data.frame(stock = x_grid, det = det$D[,1], g = g$D[,1], m = m$D[,1], gm = gm$D[,1], gi = gi$D[,1], gmi = gmi$D[,1]), id = "stock")
ggplot(policy) + geom_point(aes(stock, stock-x_grid[value], color=variable)) + ylab("escapement") 
ggplot(policy) + geom_smooth(aes(stock, stock-x_grid[value], color=variable))+ ylab("escapement")

value <- melt(data.frame(stock = x_grid, det=det$V g = g$V, m = m$V, gm = gm$V, gi = gi$V, gmi = gmi$V), id = "stock")
ggplot(value) + geom_point(aes(stock, value, color=variable)) + ylab("Net Present Value")
ggplot(value) + geom_smooth(aes(stock, value, color=variable)) + ylab("Net Present Value")
````

## Simulations

``` {r } 
simulatereps <- function(opt, true_g, true_m, true_i){
  z_g <- function() rlnorm(1,  0, true_g) 
  z_m <- function() rlnorm(1,  0, true_m) 
  z_i <- function() rlnorm(1,  0, true_i)

  sims <- lapply(1:100, function(i){
    ForwardSimulate(f, pars, x_grid, h_grid, x0 = K, opt$D, z_g, z_m, z_i, profit)
  })

  list(sims = sims, opt = opt, policy_uncertainty = c(policy_g, policy_m, policy_i),
  true_stochasticity = c(true_g, true_m, true_i))
}
````

Corner cases

```{r }
base <- simulatereps(det, 0, 0, 0)
full <- simulatereps(gmi, .2, .2, .2)
cautious <- simulatereps(gmi, 0, 0, 0)
reckless <- simulatereps(det, .2, .2, .2)
````
